---
title: "Classification iiia"
author: "Anna Heltz"
date: "2022-11-10"
output: html_document
---

```{r, include = FALSE }
library(tidyverse)
library(caret)
```

```{r}
DF <- readr::read_csv("fall2022_finalproject.csv", col_names = TRUE)
```
```{r}
DF %>% glimpse()
```


-------------------------
creating the DF in this file
----------------------------

```{r}
newDF <- DF %>% 
  mutate(x5 = 1 - (x1 + x2 + x3 + x4),
         w = x2 / (x3 + x4),
         z = (x1 + x2) / (x5 + x4),
         t = v1 * v2,
         outcome = ifelse(output < 0.33, 'event', 'non_event'),
         outcome = factor(outcome, levels = c("event", "non_event")),
          y = boot::logit(output),
         y1 = ifelse(outcome=="event",1,0))  
```

```{r}
newDF %>% glimpse()
```




--------------------------------------
Let's fit 3 models with the base features
---------------------------------------
x5 is not a base feature

All linear additive features:
```{r}
additiveBaseModel <- glm(y1~x1+x2+x3+x4+v1+v2+v3+v4+v5, data = newDF)
additiveBaseModel
```

Interaction of the categorical input with all continuous inputs:
```{r}
categoricalContinuousBaseModel <- glm(y1~x1:m+x2:m+x3:m+x4:m+v1:m+v2:m+v3:m+v4:m+v5:m, data = newDF)
categoricalContinuousBaseModel
```

All pair-wise interactions of the continuous inputs:
```{r}
pairwiseContinuousBase <- glm(y1 ~ (x1+x2+x3+x4+v1+v2+v3+v4+v5)^2, data=newDF)
pairwiseContinuousBase
```

--------------------------------------------------------------
3 Models using the “expanded feature” set
--------------------------------------------------------------
All linear additive features:
```{r}
additiveBaseDerived <- glm(y1~x5+w+z+t+v1+v2+v3+v4+v5, data = newDF)
additiveBaseDerived
```

Interaction of the categorical input with all continuous inputs:
```{r}
categoricalContinuousDerived <- glm(y1~x5:m+w:m+z:m+t:m+v1:m+v2:m+v3:m+v4:m+v5:m, data = newDF)
categoricalContinuousDerived
```
All pair-wise interactions of the continuous inputs:
```{r}
pairwiseContinuousDerived <- glm(y1 ~ (x5+w+z+t+v1+v2+v3+v4+v5)^2, data=newDF)
pairwiseContinuousDerived
```

-------------------------------------------------------------
3 Models linear basis function models:
-------------------------------------------------------------

```{r}
linearBasisMod1 <- glm(y1 ~ splines::ns(x1, 5) * splines::ns(x2, 5), data = newDF)
linearBasisMod1
```
```{r}
linearBasisMod2 <- glm(y1 ~ splines::ns(x1, 5) * (I(x2) + I(x2^2)) + t + m, data = newDF)
linearBasisMod2
```
```{r}
linearBasisMod3 <-  glm(y1 ~ splines::ns(x1, 5) * (I(x2^2)) * (I(x3^2))+ t + z + w , data = newDF)
linearBasisMod3
```
```{r}
linearBasisMod4 <-  glm(y1 ~ splines::ns(x1, 5) * splines::ns(z, 5)+ m , data = newDF)
linearBasisMod4
```



--------------------------------------------------------------------------------------
**Lets find out which model is the best based on AIC,  and BIC!!**
--------------------------------------------------------------------------------------

```{r}
extract_metrics <- function(mod, mod_name)
{
  broom::glance(mod) %>% mutate(mod_name = mod_name)
}

all_metrics <- purrr::map2_dfr(list(additiveBaseModel,additiveBaseDerived, categoricalContinuousBaseModel, categoricalContinuousDerived, pairwiseContinuousBase, pairwiseContinuousDerived, linearBasisMod1,linearBasisMod2,linearBasisMod3,linearBasisMod4),
                               as.numeric(1:10),
                               extract_metrics)
```

```{r}

all_metrics %>% 
  select(mod_name, AIC, BIC) %>% 
  pivot_longer(!c("mod_name")) %>% 
  ggplot(mapping = aes(x = mod_name, y = value)) +
  geom_point(size = 5) +
  facet_wrap(~name, scales = "free_y") +
  scale_x_continuous( breaks = c(1:10)) +
  theme_bw()
```



```{r}
AIC <- c(broom::glance(additiveBaseModel)[[4]],broom::glance(additiveBaseDerived)[[4]],broom::glance(categoricalContinuousBaseModel)[[4]],broom::glance(categoricalContinuousDerived)[[4]],broom::glance(pairwiseContinuousBase)[[4]],broom::glance(pairwiseContinuousDerived)[[4]],broom::glance(linearBasisMod1)[[4]],broom::glance(linearBasisMod2)[[4]],broom::glance(linearBasisMod3)[[4]],broom::glance(linearBasisMod4)[[4]])
AIC
BIC <- c(broom::glance(additiveBaseModel)[[5]],broom::glance(additiveBaseDerived)[[5]],broom::glance(categoricalContinuousBaseModel)[[5]],broom::glance(categoricalContinuousDerived)[[5]],broom::glance(pairwiseContinuousBase)[[5]],broom::glance(pairwiseContinuousDerived)[[5]],broom::glance(linearBasisMod1)[[5]],broom::glance(linearBasisMod2)[[5]],broom::glance(linearBasisMod3)[[5]],broom::glance(linearBasisMod4)[[5]])
BIC
cat("\nThe best model according to AIC is:", min(AIC))
cat(" which is model", which.min(AIC))
cat("\nThe best model according to BIC is:", min(BIC))
cat(" which is model", which.min(BIC))
```
As you can see model 9, or my linearBasisMod3, model is the best model.
```{r}
AIC <- AIC[-9]
BIC <- BIC[-9]
cat("\nThe second best model according to AIC is:", min(AIC))
cat(" which is model", which.min(AIC))
cat("\nThe second best model according to BIC is:", min(BIC))
cat(" which is model", which.min(BIC))
```

As you can see model 9, since we took out model 9 in the previous step, or my linearBasisMod4, model is the second best model.
```{r}
AIC <- AIC[-9]
BIC <- BIC[-9]
cat("\nThe third best model according to AIC is:", min(AIC))
cat(" which is model", which.min(AIC))
cat("\nThe third best model according to BIC is:", min(BIC))
cat(" which is model", which.min(BIC))
```

As you can see model 7, or my linearBasisMod1, model is the third best model.




-------------------------------------------------------------------
Coefficient Summaries for the top three models:
-----------------------------------------------------------------



```{r}
coefplot::coefplot(linearBasisMod1)
```
```{r}
coefplot::coefplot(linearBasisMod3)
```
```{r}
coefplot::coefplot(linearBasisMod4)
```
*How do the coefficient plots compare to the top three models:*
The coefficient plots are very similar between the top three models. The majority of coefficient values seem fairly insignificant for all 3 models, a sign that they may too be complex or possibly over fitting the data set. However, in both linear basis models 3 and 4 it seems like the coefficients for x1 splines interacting with the x2 quadratic term are very positively significant.  



